---
title: "Data Exploration Assignment-Data Analysis"
author: Melinda Davis
format: html
editor: visual
---

## Import Data from Data Preparation

Appropriate libraries and cleaned data are loaded.

```{r}
# Load Libraries
library(rio)
install_formats()
library(tidyverse)
library(stringr)
library(lubridate)
library(dplyr)
library(fixest)
library(vtable)
# Load cleaned data
cleaned_data <- rio::import("cleaned_data.csv")
```

## Addressing college interest as it relates to high-earning colleges

In September 2015, the College Scorecard became available to the public. The Scorecard released a wealth of information about colleges including incomes of graduates, SAT and ACT scores of incoming students, and the types of degrees awarded. The current question proposed is: did the release of the Scorecard shift student interest of high-earning colleges relative to low-earning ones? To start to understand this relationship, a scatter plot was created investigating the relationship of the `mean_standardized_index` (a measure of Google searches used as a proxy for interest in colleges) with time. A green line designates before and after the release of the College Scorecard. The first thing to note when examining this relationship is that there appears to be a slight decrease in search interest of all schools over time. However, high-earning versus low-earning schools shows that low-earning schools experienced a larger decrease in search interest (as can be observed by the slope of the trend-line). While this graph gives insight into what we might see in a regression analysis, it is flawed in that it shows the trend-line over time. To appropriately answer this question, a regression should be performed using dummy variables for Scorecard and high-earning colleges.

```{r}
ggplot(cleaned_data, aes(x = month, y = mean_standardized_index, color = HighEarner)) +
  geom_jitter() +
  geom_smooth(data = cleaned_data %>% filter(HighEarner), method = 'lm', se = FALSE, color = 'red') +
  geom_smooth(data = cleaned_data %>% filter(!HighEarner), method = 'lm', se = FALSE, color = 'blue') +
  geom_vline(xintercept = as.numeric(as.Date("2015-09-01")), linetype = "dashed", color = "green") +
  geom_text(aes(x = as.Date("2015-09-01"), y = max(cleaned_data$mean_standardized_index), label = "Scorecard Release 09-2015"), 
            color = "green", hjust = 0.5, vjust = -0.5) +
  geom_text(data = cleaned_data %>% filter(HighEarner), aes(label = "High Earner Trend"), x = max(cleaned_data$month), y = max(cleaned_data$mean_standardized_index), color = 'red', hjust = 3, vjust = -0.2) +
  geom_text(data = cleaned_data %>% filter(!HighEarner), aes(label = "Low Earner Trend"), x = max(cleaned_data$month), y = min(cleaned_data$mean_standardized_index), color = 'blue', hjust = 3, vjust = -0.4) +
  labs(title = "Interest in Colleges over Time",
       x = "Month",
       y = "Mean Standardized Index") +
  theme_minimal()


```

## OLS Model

To further examine this question, a fixed-effects model was performed where `mean_standardized_index` is the dependent variable and `HighEarner` and `ScoreCard` individually and as a interaction are the dependent variables. `mean_standardized_index` is a standardized proxy for school interest. `HighEarner` and `ScoreCard` are dummy variables, where 1 is assigned to these values if they are TRUE. The interaction term is included to capture the interactive effect of the two independent variables.

```{r}
# Regression Models regressing google searches on ScoreCard and highearners
model <- feols(mean_standardized_index ~ HighEarner + ScoreCard + HighEarner * ScoreCard, data = cleaned_data)

```

## Regression Analysis

The `etable(model)` shows an intercept of this model is 0.0474, which indicates the search interest of low-earning colleges before the Scorecard was introduced (when all other coefficients are 0). It is highly significant to 0.1%. `HighEarner` variable has a coefficient of -0.0245. With a p-value of greater than 5%, it is not statistically significant. The `ScoreCard` variable has a coefficient of -0.2146 and is considered highly significant. This implies that the introduction of the College Scorecard has a negative effect that is significant on the mean_standardized_index. Finally, the interaction variable `HighEarner x ScoreCard` has a coefficient of 0.1223 and is considered moderately significant to 1%. The interaction between being a `HighEarnerTRUE` and the `ScorecardTRUE` has a moderate positive effect on the `mean_standardized_index`. The adjusted R-squared is relatively low (0.02118) indicating there might be factors in the error term that were not included in the model that could potentially influence the mean_standardized_index. However, none of these variables were determined to be a source of omitted variable bias and were therefore excluded from this model.

In conclusion, the introduction of the College Scorecard increased search activity on Google Trends for colleges with high-earning graduates by 0.1223 units relative to what it did for colleges with low-earning graduates, with a standard error of 0.0420. The result comes from the HighEarnerTRUE x ScoreCardTRUE coefficient in my regression.

```{r}
etable(model)
summary(model)
```
